The Atala PRISM Protocol

### 2022-12-22 prism-backend/research/prism-protocol/did_atala.tex
Structuring proof strategy.

\commentwho{Jesus}{What follows is essentially right on its own, but I interpreted
  \cite{bhz21} somehow wrongly. The approach is: (1) merge all (ideal) global
  subroutines (ledger, RO, clock) into one (\gAtala); (2) prove that theorem 1
  below is met, where the real protocol is \RealPKIDIDAtala, and the ideal
  (global) functionality is \IdealGPKIDID -- although we can ignore that it is
  a global functionality here. Then, it would be nice to prove that Cardano as
  instantiation of the ledger, and whatever else for RO and clock, allow to
  instantiate the merged global subroutine (this is what I've mostly done next,
  to some extent). So, basically, revert the order of what I was trying to do.}

@@

At a high level, to apply the results in \cite{bch+20,bhz21}, we show the global
subroutines used in \IdealGPKIDID and \RealPKIDIDAtala can be merged into the
single ``meta'' global subroutine $\hat{\gamma}_{\gamma_1,\dots,\gamma_n}$
required by \thmref{thm:uc-bhz21}, which furthermore meets the conditions to be
subsequently suitable for UC-emulation. In addition, we argue that this is
usually the case for the known realisations of each respective global
subroutine. (This suggests that our approach is natural, and indeed
UC-realisable in practice.)
%
We then describe how the management protocol \UCM works in conjunction with
our \IdealGPKIDID and the ``meta'' global subroutines in the ideal world (i.e.,
$\UCM[\IdealGPKIDID,\hat{\gamma}_{\gamma_1,\dots,\gamma_n}]$), and
with \RealPKIDIDAtala and the ``meta'' global subroutine in the real world
(i.e., $\UCM[\RealPKIDIDAtala,\hat{\gamma}_{\gamma_1,\dots,\gamma_n}]$).
%
Finally, we show that the latter emulates the former. This concludes the proof
for \thmref{thm:uc-bhz21}.
%
We emphasise that this is merely an application of the results in
\cite{bch+20,bhz21} to our setting. The previously sketched process is detailed
next, directly referencing the relevant theorems of the corresponding works when
appropriate.

>>>>

We now prove that \RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID, in the
presence of global subroutines \IdealGclock, \IdealGRO, and \IdealGdledger.
Given that (multiple) global subroutines are involved, the proving process
requires several steps, which we summarise first, and then specify in detail. In
this process, we rely on recent results in \cite{bch+20,bhz21} as well as,
naturally, on the UC framework \cite{canetti00}. In a nutshell, the steps we
take are as follows (here, we are intentionally imprecise on occasions, in
favour of giving a high-level idea of the overall approach):

\todo{Some diagram would probably help much here.}
\begin{enumerate}
\item \emph{(Merging global subroutines into \gAtala)} Following the strategy
  in \cite{bhz21}, we combine all the global subroutines needed in our protocols
  into a single (equivalent) one. While merely ``bureaucratic'', this
  facilitates the analysis in subsequent steps. We refer to this combined
  global shared subroutine as \gAtala.
\item \emph{(\gAtala is subroutine respecting)} We show that only main ITIs
  in \gAtala accept input from ITIs not in \gAtala, no ITI within \gAtala passes
  input to ITIs not in \gAtala, no ITI in \gAtala expects subroutine output from
  other ITIs outside of \gAtala, and all subroutine outputs produced within
  \gAtala are sent to the outside world via \gAtala's main ITIs.
\item \emph{(\RealPKIDIDAtala is \gAtala-subroutine respecting)} We show that
  \RealPKIDIDAtala is also subroutine respecting, as just described for \gAtala,
  except actually for calls to \gAtala. This means that ITIs within
  \RealPKIDIDAtala may communicate directly with machines in \gAtala without
  needing to ``pass through'' the main ITIs of \RealPKIDIDAtala, and can
  communicate with them even if they are not part of \RealPKIDIDAtala's session.
\item \emph{(\IdealGPKIDID is \gAtala-subroutine respecting)} Same as before,
  but with \IdealGPKIDID.
\item \emph{(\gAtala is \RealPKIDIDAtala-regular)} We show that ITIs in \gAtala
  do not invoke new ITIs of \RealPKIDIDAtala as subroutines, nor call or expect
  subroutine output from machines of \RealPKIDIDAtala.
\item \emph{(\gAtala is \IdealGPKIDID-regular)} Same as before, but with
  \IdealGPKIDID.
\item \emph{(Executions of $\UCM[\RealPKIDIDAtala,\gAtala]$ are indistinguishable
    from execution of \RealPKIDIDAtala in the presence of \gAtala)} Given the
  previous preconditions, \cite[Proposition 3.4]{bch+20} ensures that we can
  safely apply a transformation $\UCM[\cdot,\cdot]$ to \RealPKIDIDAtala, which
  allows to leverage the techniques from the plain UC framework.
\item \emph{(Executions of $\UCM[\IdealGPKIDID,\gAtala]$ are indistinguishable
    from execution of \IdealGPKIDID in the presence of \gAtala)} Same as before,
  but with \IdealGPKIDID.
\item \emph{(\RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID in the presence
    of \gAtala)} By \cite[Definition 3.1]{bch+20}, if $\UCM[\RealPKIDIDAtala,
  \gAtala]$ $\xi$-UC-emulates $\UCM[\IdealGPKIDID,\gAtala]$, then
  \RealPKIDIDAtala $\xi$-UC-emulates \IdealGPKIDID, in the presence
  of \gAtala.
\end{enumerate}

We now proceed to specify in more detail the previous steps.

\paragraph{1. Merging global subroutines into \gAtala.} This transformation
is straightforward, and can be done as in \cite[Definition 4.1]{bhz21}. Namely,
let $\gamma_1 \gets \IdealGledger$, $\gamma_2 \gets \IdealGRO$, and $\gamma_3
\gets \IdealGclock$. Then, we define \gAtala to be a functionality that simply
accepts inputs of type $(\uccmd{query},\sid,i,x)$, and makes a query to
$\gamma_i$, with input $x$ on session \sid -- e.g., a query $(\uccmd{query},
\sid,2,(\uccmd{HasQuery},m))$ results in a query $(\uccmd{HashQuery},\sid,m)$ to
\IdealGRO.
%
This allows us to directly apply the results in \cite{bch+20}. \todo{Later, we
  will get back to arguing that the usual candidates for UC-realising the
  individual $\gamma_i$'s meet certain properties that make them suitable for
  UC-realising the ``merged'' functionality \gAtala. Thus, our main results
  apply when considering real instantiations of the global subroutines, rather
  than their idealised counterparts.}

\paragraph{2. \gAtala is subroutine respecting.}

\commentwho{Jesus}{Not removing what follows for now,  because it will be
  useful. But keep on the proof process stated in the previous enumeration.}

@@

At a high level, to apply the results in \cite{bch+20,bhz21}, we merge the
global subroutines used in \IdealGPKIDID and \RealPKIDIDAtala into a single
``meta'' global subroutine $\gAtala = \hat{\gamma}_{\gamma_1,\dots,\gamma_n}$
required by \thmref{thm:uc-bhz21}.
%
Then, to prove UC emulation, we follow the approach in \cite{bch+20} and
describe a management protocol \UCM

which furthermore meets the conditions to be
subsequently suitable for UC-emulation. In addition, we later argue that this is
usually the case for the known realisations of each respective global
subroutine. (This suggests that our approach is natural, and indeed
UC-realisable in practice.)
%
We then describe how the management protocol \UCM works in conjunction with
our \IdealGPKIDID and the ``meta'' global subroutines in the ideal world (i.e.,
$\UCM[\IdealGPKIDID,\hat{\gamma}_{\gamma_1,\dots,\gamma_n}]$), and
with \RealPKIDIDAtala and the ``meta'' global subroutine in the real world
(i.e., $\UCM[\RealPKIDIDAtala,\hat{\gamma}_{\gamma_1,\dots,\gamma_n}]$).
%
Finally, we show that the latter emulates the former. This concludes the proof
for \thmref{thm:uc-bhz21}.
%
We emphasise that this is merely an application of the results in
\cite{bch+20,bhz21} to our setting. The previously sketched process is detailed
next, directly referencing the relevant theorems of the corresponding works when
appropriate.
